Spreading of a Facebook Application

### 2011-04-15 paper.tex
checked abstract to results - metrics

@@

\subsubsection{Metrics collected by the tool}
As detailed in section \ref{sec:data-evaluation}, the tool allows for
output of metrics as \ac{csv}. Metrics about users who participated in
the quiz are given in table
\ref{tab:spread}. Table \ref{tab:nouser}
describes user who did not participate. Table \ref{tab:scorespread}
and \ref{tab:nospread} sum up metrics from table \ref{tab:spread}.

Five users participated in the application, but

Two tables also located in the addendum describe to how many
other participants friendships existed (see table \ref{tab:spread} and
\ref{tab:nospread}), and to how many people of a
\textit{new} network the application was (probably) distributed by the
user. $26$ participants were on average $2.54$ friends with other
participants and those, who spread the application, seemed to have
more connections to other people than those who did not.
The average score of people connected to others was $3.02$. It was
higher than the average score of people not connected to others ($2.41$).
People not connected to further participants had fewer friends on average ($106.8$)
than the connected participants ($138.83$).

Table \ref{tab:nouser} details how many friends who participated in
the application user had, who did not participate in the
application. Most times, those people just had one friend, but in eleven
cases, they had four or five participants in their friends list.

\subsection{Spreading of the quiz}
\label{sec:distribution-quiz}
In this section, graphs illustrating how the
application spread are presented. The original user names collected from Facebook were made anonymous
before the graphs were printed (see section \ref{sec:data-evaluation}). Friendships are
represented by directed edges: If user A has user B in the friend list, an
arrow exists targeting user B.

Figure \ref{fig:allUsers} represents each user stored in user
table as node, and friendships of those who participated in the
application as directed edges. Names of users who participated in the quiz are distinguished by
red font color. The main purpose of this figure is to visualize the
whole setting and to give an imagination of how many \textit{new}
people on Facebook could have read about the application in their 
news stream. In addition, the figure shows few participants, who are
not connected to the larger network of the 
person who started to spread the application.

In the following, only user who participated in the quiz are
considered. In Figure \ref{fig:spreading} nodes of people, who are
connector between two networks, have a red font color. The number of
friends and the average score from quiz are also shown. 
It can be obtained that in three cases, there are several connections (friendships)
between the people who participated and belong to a new network. For
instance, several connections exist between users \textit{Evan},
\textit{Diego}, \textit{Isabelle}, \textit{Chase}, and \textit{Gebrielle}.
\textit{Nicholas} and \textit{Chase}, who are friends with those
people mentioned above, also are friends with \textit{Ethan},
and \textit{Josiah}.  Then, \textit{Faith} is friend with
\textit{Owen} and \textit{Colton}, who also know each other, and
\textit{Mariah} and \textit{Sydney} have \textit{Alexis} as a common
friend. The figure also shows that most people who spread the
application had a score greater than or equal to three. At least, it
can be obtained that half of the people who spread the application
had less than hundred friends, which is less than the mean on
Facebook, i.e. 130 friends \cite{facebookStats}.

In the addendum (section \ref{sec:addendum}), two further graphs are
shown. In figure \ref{fig:friendsScored}, the background color
represents a user's average score. User with a darker background had a
higher score with those having a lighter background color. 
Figure \ref{fig:friendsFilled} visualizes the amount of friends in the
by giving users having many friends a dark and those having fewer
friends a lighter background color.

@@

\section{Discussion}
\label{sec:discussion}
From the results given in section \ref{sec:results} it can be derived
that the number of friends did not to play a bigger role in the
process of the application's distribution. The number of connections
to other users seems to play a bigger 
role, as it is indicated by table \ref{tab:spread} and figure
\ref{fig:spreaders}. In addition, a worse score made people unlikely
to spread the application, just one person with a score less than 3
did so. Thus, people with a higher score (on average $3.11$, median $3.32$) are
supposed to have distributed the application. Indeed, some people who
had four or five friends who participated in the application did not
install the application.

\section{Conclusion and Outlook}
\label{sec:conclusion}
This research indicates, that the spreading of application follows
certain patterns. Oftentimes, people say that they don't want to
make a worse impression on Facebook, and this is also reflected in the
fact, that only people who yield higher scores in the quiz also
distributed to their friends. In the environment analyzed, dense
networks with few participants, i.e. in networks in which people know
each other, made  further members of their network very likely to join
the application. Indeed, the number of friends did not play a role in
making people likely to spread the application.

@@

\begin{figure}
  \includegraphics[height=3in]{../img/newNW.png}
 \caption{This figure emphasizes users who connect to new social networks with
   red border.}
\label{fig:newNW}
\end{figure}

@@

 \begin{table}
 \label{tab:spread}
 \begin{tabular}{ c c c c c}
Anon. Name&Friends&App-friends&Spread App to&Average Score	\\
Alexandra & 133& 34& 34 & 4.4625 \\
Evan & 97& 1& 0 & 3.33333 \\
Justin & 90& 2& 0 & 1.28571 \\
Tristan & 332& 2& 0 & 3 \\
Thomas & 396& 3& 0 & 4 \\
Jose & 233& 1& 0 & 4 \\
David & 322& 6& 0 & 1.5 \\
Gianna & 63& 3& 1 & 4 \\
Makayla & 79& 1& 0 & 3 \\
Connor & 102& 3& 0 & 0 \\
Ethan & 91& 4& 1 & 1 \\
Isabelle & 53& 1& 0 & 1 \\
Gavin & 89& 1& 0 & 2 \\
Caleb & 128& 3& 0 & 4 \\
Sarah & 151& 5& 3 & 3 \\
Emma & 72& 1& 0 & 4 \\
Gabrielle & 135& 4& 2 & 3.63636 \\
David & 254& 3& 2 & 4 \\
Emily & 220& 5& 0 & 3 \\
Victoria & 63& 2& 0 & 2 \\
Brian & 126& 2& 0 & 4 \\
Jose & 127& 3& 0 & 3 \\
Thomas & 186& 4& 0 & 4 \\
Evelyn & 42& 5& 0 & 3 \\
Paige & 76& 2& 0 & 1 \\
Hannah & 264& 2& 0 & 1 \\
Charlotte & 445& 5& 0 & 3 \\
Elizabeth & 35& 1& 0 & 3 \\
Carson & 153& 3& 1 & 3 \\
Carson & 20& 2& 1 & 3 \\
Nathan & 121& 1& 0 & 4 \\
Eli & 11& 9& 2 & 4.23077 \\
Isaac & 290& 1& 0 & 1 \\
Alexandra & 130& 1& 0 & 3 \\
Gabriel & 67& 1& 0 & 2 \\
Paige & 16& 1& 0 & 4 \\
David & 50& 1& 0 & 4 \\
Julian & 207& 2& 0 & 2 \\
Carlos & 3& 3& 0 & 4.75 \\
Evan & 74& 3& 0 & 4.5 \\
Sebastian & 47& 3& 0 & 4.41667 \\
Xavier & 80& 0& 0 & 4 \\
Evan & 157& 2& 0 & 3 \\
Brody & 136& 2& 0 & 3 \\
Luis & 60& 1& 0 & \\
Allison & 13& 0& 0 & 2.5 \\
Tyler & 229& 0& 0 & \\
Emily & 53& 0& 0 & \\
Khloe & 78& 1& 0 & 4 \\
Katelyn & 167& 1& 0 & 2 \\
Kaitlyn & 10& 1& 0 & 3 \\
Eva & 272& 0& 0 & 4
\end{tabular}
\caption{Table details metrics about users who participated in the
  application. Friends detail the number of friends a user had,
  App-friends is the number of friends who also participated in the
  quiz, Spread App To details to how many people a user probably
  spread the application, and Average Score details to the score the
  user averagely had in the quiz. The names were made anonymous.}
\end{table}

@@

\begin{table}
\label{tab:scorespread}
\begin{tabular}{c c}
Spread to & Average score \\
1&	2.75 \\
2&	3.96 \\
3&	3 \\
\end{tabular}
\caption{Table details the average score of people who spread the
  application to one, two, or three users.}
\end{table}

>>>>

\subsection{Start of the spreading process}
\label{sec:start-spre-proc}
The spreading process was started by the author who posted in the news
stream a German analogon to `please check out my new app!'. In
addition, a message on Facebook was sent to 17 friends asking them to
participate.

@@

\subsubsection{Metrics collected by the tool}
\label{sec:tool-metrics}
As detailed in section \ref{sec:data-evaluation}, the tool allows for
output of metrics as \ac{csv}. Metrics about users who participated in
the quiz are given in table
\ref{tab:spread}. Table \ref{tab:nouser}
describes user who did not participate. Table \ref{tab:scorespread}
and \ref{tab:nospread} sum up metrics from table \ref{tab:spread}.

From the tables it can be derived that 34 users belonged to the
network of the author who started to spread the application. Thus, 18
application users came from other networks. 5 user did not have any
friends who also participated. On average, people except the author
and who had at least one \textit{app friend} had on average $2.5$ friends who also
used the application. People who probably spread the application to
one person had on average $3$ friends who used the quiz, too. People
who probably spread to two persons had on average $5.33$ friends who
used it, too. As it is shown in table \ref{tab:scorespread},
the average score of people who probably spread 
the application was higher for people who spread to more people. 
% Two tables also located in the addendum describe to how many
% other participants friendships existed (see table \ref{tab:spread} and
% \ref{tab:nospread}), and to how many people of a
% \textit{new} network the application was (probably) distributed by the
% user. $26$ participants were on average $2.54$ friends with other
% participants and those, who spread the application, seemed to have
% more connections to other people than those who did not.
% The average score of people connected to others was $3.02$. It was
% higher than the average score of people not connected to others ($2.41$).
% People not connected to further participants had fewer friends on average ($106.8$)
% than the connected participants ($138.83$).

Table \ref{tab:nouser} describes friends of participants, who did not
participate in the quiz. As each friend was stored, these people also
occur in table \textit{User} and \textit{Friends}. These friends of
application users may have received friend requests from other users,
and, they may have read about the application in their friends' news
stream.
It can be obtained, that 11 users exist having 4 or five friends who
use the application. The larger amount of people had fewer friends who
used the application.

\subsection{User Response}
\label{sec:user-response}
As mentioned in section \ref{sec:start-spre-proc}, 17 friends of the author were
directly invited to participate in the application. Each of them
installed the application, thus 17 friends of the author installed the
application for other reasons. The number of installations per day is
shown in figure \ref{fig:installations}, the data were exported from
Facebook Insights. A peak of new installations could be
found after the request to participate was posted in the news stream
on 3rd of March, 2011. The days after, further installations
happened. Six of the people who were asked directly to participate,
were not asked on 3rd of March, but afterwards. 7 people were asked
before.

\subsection{Graphs detailing the spreading of the quiz}
\label{sec:distribution-quiz}
In this section, graphs illustrating how the
application spread are presented. The original user names collected
from Facebook were made anonymous
 (see section \ref{sec:data-evaluation}). Friendships are
represented by directed edges: If user A has user B in the friend list, an
arrow exists targeting user B.

Figure \ref{fig:allUsers} represents the graph that is created when
parameter \verb|tree| is set. In this case, each user stored in \textit{User}
table becomes a node, and friendships stored in table \textit{Friends}
are represented by directed edges. Names of users who participated in the quiz are distinguished by
red font color. The main purpose of this graph type is to
visualization of the whole spreading process.
The graph gives an imagination of how many 
people on Facebook could have read about the application in their 
news stream, or could have been invited by friends to join the
application. In addition, the graph shows how new networks are
reached and how dense the connections were in the network of the root
of the spreading process.

Figure \ref{fig:spreading} shows the graph that can be obtained when
parameter \verb|spreading| is set. In this case, only users who
participated in the quiz are considered. It can be obtained, that the
users Evelyn, Elli, Carlos, Charlotte, David, Sebastian,
Evan, Thomas, and Gabrielle have several connections among themselves,
thus they belong probably to a social network. Most people of this
network had a relatively high score. 
Further smaller networks exist. One among the users Sarah, two Evans,
and Brody, and another one among Julian, Khole, David, and Carson. In
these networks, the application spread to people of a new network. In
other cases, one new person who had a friend who used the application
started to use it as well. In
addition, some further people found the application in the application
directory on Facebook.
At least, it can be obtained that half of the people who spread the application
had less than hundred friends, which is less than the mean on
Facebook, i.e. 130 friends \cite{facebookStats}.

% In the addendum (section \ref{sec:addendum}), two further graphs are
% shown. In figure \ref{fig:friendsScored}, the background color
% represents a user's average score. User with a darker background had a
% higher score with those having a lighter background color. 
% Figure \ref{fig:friendsFilled} visualizes the amount of friends in the
% by giving users having many friends a dark and those having fewer
% friends a lighter background color.

@@

\section{Discussion}
\label{sec:discussion}
From the results given in section \ref{sec:results} it can be followed
that mainly people belonging to the author's social network
participated in the application. Many of them came after the
application was solicited on Facebook's news stream, but in the days
after, further but smaller numbers of installations happened. As participation of friends was posted
in the news streams of the participants, friends could read about the
quiz few times. This may have made people curious about the application.
As it can be obtained in the graphs and in the
metrics, some users did not have friends who participated. Thus, they
came for another reason, may be due to interest in dentistry.

Most people who participated were in the age of twenty five to thirty
four. One reason may be that most of the author's friends are of that
age. As teenagers have less teeth affected with
caries than people older than twenty \cite{zahngesundheit}, the chosen
topic of the question can have increased the effect that mainly people
in the twenties and thirties participated.

From the metrics it can be concluded that the number of friends did
not to play a bigger role in the 
process of the application's spreading. The number of connections
to other users and the quiz score seemed to play a bigger 
role, as it is mentioned in section \ref{sec:tool-metrics}. 
In addition, a worse score made people unlikely
to spread the application, just one person with a score less than 3
did so. Thus, people with a higher score (on average $3.11$, median $3.32$) are
supposed to have distributed the application. Indeed, some people who
had four or five friends who participated in the application did not
install the application.

\section{Conclusion and Outlook}
\label{sec:conclusion}
This research indicates, that their is not just one default pattern
after which applications spread on Facebook. Instead, several facts
seem to influence this effect. On the one hand, the topic of an
application and an applications type hit the interest of specific
people. Furthermore, there are often

. Oftentimes, people say that they don't want to
make a worse impression on Facebook, and this is also reflected in the
fact, that only people who yield higher scores in the quiz also
distributed to their friends. In the environment analyzed, dense
networks with few participants, i.e. in networks in which people know
each other, made  further members of their network very likely to join
the application. Indeed, the number of friends did not play a role in
making people likely to spread the application.

@@

\begin{figure}
  \includegraphics[height=1.5in]{../img/newNW.png}
 \caption{This figure emphasizes user b who connects the existing
   network of users to a new social networks with
   red border.}
\label{fig:newNW}
\end{figure}

\begin{figure}
  \includegraphics[width=5in]{../img/installations.png}
\caption{Figure shows all users stored in the database. Users who
  participated have a red text color.}
\label{fig:installations}
\end{figure}

@@

 \begin{table}
 \label{tab:spread}
 \begin{tabular}{ c c c c c}
Anon. Name&Friends&App-friends&Spread App to&Average Score	\\
Alexandra & 133& 34& 34 & 4.4625 \\
Evan & 97& 1& 0 & 3.33333 \\
Justin & 90& 2& 0 & 1.28571 \\
Tristan & 332& 2& 0 & 3 \\
Thomas & 396& 3& 0 & 4 \\
Jose & 233& 1& 0 & 4 \\
David & 322& 6& 0 & 1.5 \\
Gianna & 63& 3& 1 & 4 \\
Makayla & 79& 1& 0 & 3 \\
Connor & 102& 3& 0 & 0 \\
Ethan & 91& 4& 1 & 1 \\
Isabelle & 53& 1& 0 & 1 \\
Gavin & 89& 1& 0 & 2 \\
Caleb & 128& 3& 0 & 4 \\
Sarah & 151& 5& 3 & 3 \\
Emma & 72& 1& 0 & 4 \\
Gabrielle & 135& 4& 2 & 3.63636 \\
David & 254& 3& 2 & 4 \\
Emily & 220& 5& 0 & 3 \\
Victoria & 63& 2& 0 & 2 \\
Brian & 126& 2& 0 & 4 \\
Jose & 127& 3& 0 & 3 \\
Thomas & 186& 4& 0 & 4 \\
Evelyn & 42& 5& 0 & 3 \\
Paige & 76& 2& 0 & 1 \\
Hannah & 264& 2& 0 & 1 \\
Charlotte & 445& 5& 0 & 3 \\
Elizabeth & 35& 1& 0 & 3 \\
Carson & 153& 3& 1 & 3 \\
Carson & 20& 2& 1 & 3 \\
Nathan & 121& 1& 0 & 4 \\
Eli & 11& 9& 2 & 4.23077 \\
Isaac & 290& 1& 0 & 1 \\
Alexandra & 130& 1& 0 & 3 \\
Gabriel & 67& 1& 0 & 2 \\
Paige & 16& 1& 0 & 4 \\
David & 50& 1& 0 & 4 \\
Julian & 207& 2& 0 & 2 \\
Carlos & 3 & 3 & 0 & 4.75 \\
Evan & 74& 3& 0 & 4.5 \\
Sebastian & 47& 3& 0 & 4.41667 \\
Xavier & 80& 0& 0 & 4 \\
Evan & 157& 2& 0 & 3 \\
Brody & 136& 2& 0 & 3 \\
Luis & 60& 1& 0 & 0\\
Allison & 13& 0& 0 & 2.5 \\
Tyler & 229& 0& 0 & 0\\
Emily & 53& 0& 0 & 0\\
Khloe & 78& 1& 0 & 4 \\
Katelyn & 167& 1& 0 & 2 \\
Kaitlyn & 10& 1& 0 & 3 \\
Eva & 272& 0& 0 & 4
\end{tabular}
\caption{Table details metrics about users who participated in the
  application. Friends detail the number of friends a user had,
  App-friends is the number of friends who also participated in the
  quiz, Spread App To details to how many people a user probably
  spread the application, and Average Score details to the score the
  user averagely had in the quiz. The names were made anonymous.}
\end{table}

@@

\begin{table}
\label{tab:scorespread}
\begin{tabular}{c c}
Spread to & Concerned People & Average score \\ 
1&	4 & 2.75 \\
2&	3&3.96 \\
3&	1&3 \\
\end{tabular}
\caption{Table details the average score of people who spread the
  application to one, two, or three users, as well as the number of
  concerned people.}
\end{table}
